---
title: "Question 1"
author: "Candidate No. A13582 - Akithmi Wasalathanthri"
date: "2024-03-19"
output: html_document
---
"
******************
*"Question - 1"*
******************

**************
*"Part - A"*
**************

```{r}
# "Loading the necessary libraries"
library(stats)


f <- function(x) {
  return(0.5 * exp(-abs(x)))
}
```
# "Define the probability density function (PDF) for the Metropolis-Hastings Algorithm for f(x) over the real numbers x".

```{r}
# "giving the parameters"
N <- 10000  
s <- 1  
x <- numeric(N)  
x[1] <- rnorm(1)  
```

```{r}
for (i in 2:N) {
  # "Stating a proposal for each iteration of the loop using the normal distribution"
  x_asteric <- rnorm(1, mean = x[i-1], sd = s)
  
  # "Calculate the acceptance ratio "-
  # "the ratio of the probability density function (PDF) of the proposed value x_asteric to the PDF of the current value."
  r <- f(x_asteric) / f(x[i-1])
  
  # "Then we decide of accepting or rejecting the proposal in the context of the Metropolis-Hastings algorithm"
  u <- runif(1)
  if (u < r) {
    x[i] <- x_asteric
  } else {
    x[i] <- x[i-1]
  }
}
```

```{r}
# "Define the x values for the graph"
x_values_range <- seq(min(x), max(x), length.out = 1000)
hist(x, breaks = 50, freq = FALSE, col = 'lightgreen', main = 'Metropolis-Hastings Algorithm', xlab = 'x', ylab = 'Density', xlim = c(min(x), max(x)))
# " Kernel Density Estimate line"
lines(density(x), col = 'red', lwd = 3)
legend('topright', legend = c('Histogram', 'Kernel Density Estimate'), fill = c('lightgreen', 'red'))
```


```{r}
# "Define the x numbers for the graph"
x_values_range <- seq(min(x), max(x), length.out = 1000)


hist(x, breaks = 50, freq = FALSE, col = 'lightgreen', main = 'Metropolis-Hastings Algorithm', xlab = 'x', ylab = 'Density', xlim = c(min(x), max(x)))

#" Kernel Density Estimate line"
lines(density(x), col = 'red', lwd = 3)

# "density function f(x)"
curve(0.5 * exp(-abs(x)), col = 'blue', lty = 2, lwd = 2.5, add = TRUE)
legend('topright', legend = c('Histogram', 'Kernel Density Estimate', 'Density Function f(x)'), fill = c('lightgreen', 'red', 'blue'), lty = c(0, 1, 2))

```


```{r}
# "Define the range for the graph"
x_values_range <- seq(min(x), max(x), length.out = 1000)

# "Computing the kernel density estimate using sampled data x"
kde_est <- density(x)

# "Evaluating the KDE at x_values_range"
kde_values <- approxfun(kde_est, rule = 2)(x_values_range)

# "Computing the logarithms of the kernel density estimate values"
log_kde_values <- log(kde_values)


hist(x, breaks = 50, freq = FALSE, col = 'lightgreen', main = 'Metropolis-Hastings Algorithm', xlab = 'x', ylab = 'Density', xlim = c(min(x), max(x)))

lines(x_values_range, exp(log_kde_values), col = 'red', lwd = 3)

lines(x_values_range, exp(log_f_values), col = 'blue', lty = 2, lwd = 2.5)

legend('topright', legend = c('Histogram', 'Kernel Density Estimate', 'Density Function f(x)'), fill = c('lightgreen', 'red', 'blue'), lty = c(0, 1, 2))

```


```{r}
# "Calculating the Sample Mean and Sample Standard Deviation"
sample_mean <- mean(x)
sample_standard_dev <- sd(x)

cat("Sample Mean:", sample_mean, "\n")
cat("Sample Standard Deviation:", sample_standard_dev, "\n")

```

*****************
*"#PART - (B)"*
*****************

*****************************************************************
*"Calculating the R^ for the random walk Metropolis algorithm"*
*****************************************************************

*"importing the required libraries for part b"*
```{r}
library(ggplot2)
```

*"Defining the f(x) function that was used in part a) once again"*
```{r}
f_x <- function(x) {
  return(0.5 * exp(-abs(x)))
}
```

*"Stating the functions to calculate the sample mean and sample variance of chain j"*
```{r}
samplegiven_mean_M <- function(x) {
  return(mean(x))
}

samplegiven_var_V <- function(x, M) {
  return(mean((x - M)^2))
}
```

*"Generating multiple sequences of N samples"*
```{r}
J <- 4
N <- 2000
s_values <- seq(0.001, 1, length.out = 100)
R_hat_values <- numeric(length(s_values))

for (i in 1:length(s_values)) {
  # "Generate chains"
  chains <- matrix(0, nrow = J, ncol = N)
  for (j in 1:J) {
    chains[j, 1] <- 0  
    for (k in 2:N) {
      x_asteric <- rnorm(1, mean = chains[j, k-1], sd = s_values[i])
      r_ratio <- f_x(x_asteric) / f_x(chains[j, k-1])
      u <- runif(1)
      if (u < r_ratio) {
        chains[j, k] <- x_asteric
      } else {
        chains[j, k] <- chains[j, k-1]
      }
    }
  }
  
  # "Computing the  Mj value"
  Mj <- apply(chains, 1, samplegiven_mean_M)
  
  # "Computing the  Vj value"
  Vj <- sapply(1:J, function(j) samplevar_V(chains[j,], M = Mj[j]))
  
  # "Computing to find M, B, W, and R^ values'
  M <- mean(Mj)
  B <- mean((Mj - M)^2)
  W <- mean(Vj)
  R_hat <- sqrt((B + W) / W)
  R_hat_values[i] <- R_hat
}
```

*"Plot R^ values over s values using a line graph"*
```{r}
plot(s_values, R_hat_values, type = "l", col = "red", lwd = 2, xlab = "s", ylab = "R^", main = "R^ Values over s Values")
grid()
```

end



























."